\propertitle

### 2020-03-16 replies1_ncomp.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2020-03-16T14:36:02+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimfalse
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\firstdraft}{16 March 2020}
\newcommand*{\firstpublished}{16 March 2020}
\newcommand*{\updated}{\ifarxiv***\else\today\fi}
\newcommand*{\propertitle}{Sketch of replies to N. Comp. reviews%\\{\large ***}%
}% title uses LARGE; set Large for smaller
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{}
\newcommand*{\pdfauthor}{}
\newcommand*{\headauthor}{}
\newcommand*{\reporthead}{}%\ifarxiv\else Open Science Framework \href{https://doi.org/10.31219/osf.io/***}{\textsc{doi}:10.31219/osf.io/***}\fi}% Report number

@@

Let us emphasize, first of all, that the purpose of our paper is to make
inferences about the total activity of the larger population, not to decide
among different sufficient statistics or correlation orders. Section~4 is
meant to show the implications of our method for that kind of discussions.
In particular we numerically show that conclusions based on a sample and
conclusions based on the larger population can be very different. Our
method can bypass that difficulty, if one wants to use it for such
comparisons.

There are many differing views on the questions of comparison, complexity,
nesting, and \enquote{penalization} of models. Our view is close to that
expressed in several works, studies, and commentaries
\citep[\eg][]{good1985,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
and well-explained in MacKay \citey{mackay1992}. We summarize this view
below. According to it, the analysis in our \sect~4, based on the weight of
evidence, is consistent and does not need any corrections.

Note in particular that we absolutely avoid speaking about
\enquote{significance}, also as recommended by the American Statistical
Association in their recent official statements \citep{asa2016}[especially
\sect~2]{asa2019}. We are simply calculating the evidence \citep{good1950}
$\p(D \|M_{i})$ for several hypotheses $M_{i}$ based on data $D$. Evidence
is a continuous quantity and we are not setting any arbitrary thresholds.
Multiplied by a pre-data probability, this evidence gives the post-data
probability for each hypothesis. The pre-data probability will depend on
the specific case, animal, brain region, and additional knowledge. If the
evidence is $10^{35}:1$ for the first hypothesis against the second, but
our pre-data probability for the second is $10^{36}$ times that for the
first, then the second hypothesis will have higher post-data probability.
But there is still a continuum of cases, no dichotomy. No such dichotomy is
necessary.

If a \emph{choice} among two or more hypotheses has to be made for some
purpose (say, hardware algorithm implementation in some automated device),
such choice requires not only the probabilities $p(H_{i}\|D)$, but also the
gain/loss matrix for the various choices conditional on the various
hypotheses. The choice is then made by maximization of the expected gain.
Some gain matrices can even lead to choosing the hypothesis having
\emph{lower} posterior probability \citep[this frequently happens in
medical decision problems, see \eg][]{soxetal1988_r2013}.

@@

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2020-03-16T19:55:02+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourtrue% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimfalse
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\firstdraft}{16 March 2020}
\newcommand*{\firstpublished}{16 March 2020}
\newcommand*{\updated}{\ifarxiv***\else\today\fi}
\newcommand*{\propertitle}{Sketch of replies to N. Comp. reviews%\\{\large ***}%
}% title uses LARGE; set Large for smaller
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{}
\newcommand*{\pdfauthor}{}
\newcommand*{\headauthor}{}
\newcommand*{\reporthead}{}%\ifarxiv\else Open Science Framework \href{https://doi.org/10.31219/osf.io/***}{\textsc{doi}:10.31219/osf.io/***}\fi}% Report number

@@

Let us emphasize, first of all, that the purpose of our paper is to make
inferences about the total activity of the larger population, not to decide
among different sufficient statistics or correlation orders. Section~4 is
meant to show the implications of our method for that kind of discussions.
In particular we want to show that \emph{conclusions based on a sample and
  conclusions based on the larger population can be very different}. Our
method can bypass that difficulty, if one wants to use it for such
comparisons.

We partly agree with the reviewer's concern, but for different reasons. The
simple comparison of the log-evidences is not a problem in itself, as
explained below. The problem is what is meant by \enquote{model} and what
is being compared. Under one interpretation, our calculation of the
log-evidence is missing a factor, which comes from the full Bayesian
analysis. This factor, however, should be roughly the same in the
calculations from the sample and from the larger population, so our point
that the two calculations lead to different conclusions still holds. Under
another interpretation our calculation is correct as it is. We explain this
below.

In any case, please note that we avoid speaking about
\enquote{significance}, also as recommended by the American Statistical
Association in their recent official statements \citep{asa2016}[see
especially \sect~2]{asa2019}. When comparing hypotheses $M_{i}$ we simply
want their post-data probabilities $\p(M_{i} \| D)$, with their continuum
of cases. Other researchers are of course free to make dichotomies if they
like; but we personally agree with the American Statistical Association
that any threshold is unnecessary and arbitrary. If a \emph{choice} among
the hypotheses has to be made for some purpose (say, hardware algorithm
implementation in some automated device), such choice requires not only the
probabilities $\p(M_{i}\|D)$, but also the utility/loss matrix for the
various choices conditional on the various hypotheses
\citep{kadaneetal1980b}. The choice is then made by maximization of the
expected utility, a procedure dictated by some simple rational desiderata
\citep{fishburn1981,bernardoetal1994_r2000}[\chap~13]{jaynes1994_r2003}.
Some utility matrices can even lead to choosing the hypothesis having
\emph{lower} post-data probability; this frequently happens in
medical decision problems, see \eg\ \textcite{soxetal1988_r2013}.

% We want a continuous quantity, without any arbitrary thresholds. The
% evidence is such a continuous quantity. Multiplied by a pre-data
% probability, the evidence gives the post-data probability for each
% hypothesis. The pre-data probability will depend on the specific case,
% animal, brain region, and additional knowledge. If the evidence is
% $10^{35}:1$ for the first hypothesis against the second, but our pre-data
% probability for the second is $10^{36}$ times that for the first, then the
% second hypothesis will have higher post-data probability. But there is
% still a continuum of cases, no dichotomy. No such dichotomy is necessary.
% simply calculating the evidence \citep{good1950} $\p(D \|M_{i})$ for
% several hypotheses $M_{i}$ based on data $D$. Evidence is a continuous
% quantity and we are not setting any arbitrary thresholds. Multiplied by a
% pre-data probability, this evidence gives the post-data probability for
% each hypothesis. The pre-data probability will depend on the specific case,
% animal, brain region, and additional knowledge. If the evidence is
% $10^{35}:1$ for the first hypothesis against the second, but our pre-data
% probability for the second is $10^{36}$ times that for the first, then the
% second hypothesis will have higher post-data probability. But there is
% still a continuum of cases, no dichotomy. No such dichotomy is necessary.

\smallskip

There are many differing, fragmented views on the questions of comparison,
complexity, nesting, and \enquote{penalization} of models. Our view is
close to that expressed in several works, studies, and commentaries
\footcites[\eg][]{good1950,good1985,kadaneetal1980b,jonesetal1986,chatfield1995,copasetal1995,draper1995,spiegelhalteretal1995,raftery1995,gelmanetal1995,mackay1995_r2003,draperetal1996,hoetingetal1999,clydeetal1999,browneetal2006},
and well-explained in MacKay \citey{mackay1992}. According to this view,
which we summarize below, the log-evidence automatically takes care of
model complexity.

The main question is what we mean by \enquote{model}. There are two main
possibilities: model as a family, or model as a specific distribution.

\smallskip

\textbf{(A)} If by \enquote{model} we mean a family of probability
distributions, \eg\ (apart from normalization)
\[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
    \mu\in\RR, \sigma \in\opop{0,\infty}}\;,\]
then model comparison is ill-defined. % Each distribution in the family
% assigns a different probability, and it is not clear which should be
% chosen or how they should be combined and why.
It may happen that the optimal distribution in family $\mathcal{F}_{1}$
assigns more probability to the observed data than the optimal distribution
in family $\mathcal{F}_{2}$, and yet $\mathcal{F}_{1}$ assigns \enquote{on
  average} lower probabilities to the data than $\mathcal{F}_{2}$. It is
not clear how the two families should be compared in this case, or what
$\p(D \| \mathcal{F}_{i})$ is. A hypothesis is ill-defined, or at least
useless, if it does not allow us to assign a probability to the data. Such
a probability can be obtained if we specify a pre-data probability
distribution over the distributions in the family, \eg\
$\pf(\mu,\sigma \| I)\,\di\mu\,\di\sigma$ in the example above.
\enquote{Model} must therefore be redefined as a family of distributions
\emph{and} a pre-data distribution over such family. This redefinition has
several consequences:

First, two models are different if they have the same family but different
distributions over it. In fact they lead to different probabilities on the
data space. Two such models are also mutually exclusive,
$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use the one family
distribution \emph{and} the other family distribution at the same time.
It's either-or. This consequence makes sense from a nonparametric
perspective, where a parametric model (in the redefined sense) is just a
delta-like nonparametric distribution, concentrated on a lower-dimensional
manifold (the family above) of the full space of distributions
\citep[p.~761]{draperetal1996}.

Second, the \enquote{complexity} of a model cannot be judged from the
dimensionality of its family. Model $M_{1}$ can have a family distribution
concentrated on a very small region, with respect to model $M_{2}$ having
the same family. $M_{1}$ is thus effectively using fewer parameters than
$M_{2}$. Correspondingly the \enquote{typical set} of data to which $M_{1}$
assigns appreciable probability is smaller than the typical set of $M_{2}$:
model $M_{1}$ is less \enquote{powerful}. The effective dimensionality or
complexity of a model is thus best defined by the volume of data that is
typical under that model. Most important, it may then happen that a model
is more powerful than another even if it is defined on a lower-dimensional
family.

Third, \emph{the evidence $\p(D \| M_{i})$ of a model automatically takes
  care of its complexity}. This happens because the probability over the
data space is normalized. If $M_{1}$ has a typical-data volume larger than
$M_{2}$'s, then it must also assign lower probability to each data point in
that volume, to ensure normalization. As a result, if the observed data
$D^{*}$ lie in the typical volumes of both models, then
$\pf(D^{*} \| M_{1})< \pf(D^{*} \| M_{2})$: \emph{the evidence is
  automatically penalizing the larger model}. There is a continuous
trade-off between how typical the data is for each model and each model's
power. This is the so-called automatic Ockham razor of Bayes's theorem,
extensively explained by MacKay
\citey{mackay1992}[also][\sect~28.1]{mackay1995_r2003}. And this is the
reason why we are personally satisfied with simply comparing the evidence
for any two models.

In our calculations, however, we are using a zeroth-order Laplace
approximation for calculating the evidence in \sect~4. A first-order
approximation (containing the width of the distribution, influenced by the
typical-set volume above) would correct for the model's power. We believe
that the correction would nevertheless be the same in the calculation based
on the larger population, \eqns~(11), and on the sample, \eqns~(12),
because the data space is the same. Hence the difference in the two
approaches -- our main message -- should persist. But it would be better to
check this.

\smallskip

\textbf{(B)} If by \enquote{model} we mean a specific distribution, \eg\
\[ \exp\biggl[-\frac{(x-5)^{2}}{18}\biggr]\,\di x \;,\] then two models can
of course be directly compared: $\p(D \|M_{i}, I)$ is well-defined. In this
case the \enquote{dimensionality} of the model is undefined, however. The
distribution above could be considered as belonging to the family
$\set[\Big]{\exp\Bigl[-\frac{(x-\mu)^{2}}{18}\Bigr]}$, or
$\set[\Big]{\exp\Bigl[-\frac{(x-5)^{2}}{2\sigma^{2}}\Bigr]}$, or
$\set[\big]{\exp[f(x-5)]}$, and so on. A point in a manifold belongs to an
infinity of submanifolds.

But the property of having a minimal sufficient statistic is still
well-defined for a model in this sense, in an exchangeable context. The
model above has $\set[\big]{n, \widebar{x}, \widebar{x^{2}}}$ (or any
equivalent set) as minimal sufficient statistics, where the bar denotes the
sample average. Thus two models having different sufficient statistics can
be compared without reference to any family. Interpreted in this sense our
calculations of \sect~4 are correct.

@@

The question, in summary, is what exactly is being compared, whether
families (with accompanying distribution) or single distributions. The
literature is not at all clear. But, in either case, our calculations show
that the approaches based on a population and a sample thereof can lead to
different conclusions -- in accordance with the non-preservation of
sufficiency under sampling -- and they provide a way to bypass this
problem.

@@

As stated on p.~8, the interpretation is important because it leads to
numerically different results in first- and higher-order approximations;
Appendix~C explains where the differences appear. Also, it is important to
know whether we are calculating the frequency of past results, or the
probability for a new result, or the most probable frequency of future
results. These three quantities are very different when the ratio
$\frac{\text{\# samples}}{\text{data-space dimension}}$ is high, and they are
to be used in different ways.

So these are not philosophical matters, it's a matter of understanding what
we are doing and stating it clearly in our work. % Unfortunately the
% neuroscience literature is very unclear on such matters.
There's also another important reason. We are generally concerned for
master's and PhD students, who often learn concepts and methods through
actual research, perusing works like the present one. Young researchers too
often end up learning incorrect or confused ideas, gathered from unclear
literature. They trust it because it's peer-reviewed. Sometimes the result,
quoting Feynman \citey{feynman1985_r1989}, is that \enquote{they don't
  learn by understanding; they learn by some other way -- by rote, or
  something. Their knowledge is so fragile!}. We owe them clarity,
precision, honesty, and at least a glimpse of the bigger picture.
